### 2016-03-22 thesis/abstract.tex
thesis completed

Supermassive black holes in the local Universe obey a surprisingly large number of scaling laws 
that involve the black hole mass and various properties of the host spheroids. 
These ``black hole mass scaling relations'' reveal a strong symbiosis between galaxies and black holes, 
define important constraints about their co-evolution through the cosmic time, 
and set the boundary conditions (at $z=0$) for theoretical models and simulations of galaxy evolution. \\

%%A careful galaxy decomposition is required to accurately measure a galaxy's spheroid properties. 
%%Recent studies have attempted structural decomposition 
%%for similar samples of galaxies with a direct measure of the black hole mass, 
%%but they have not converged to the same conclusions. 
%%This is because their models for the same galaxy were often not consistent with each other in terms of fitted components, 
%%or, even when they used the same model for the same galaxy, they obtained significantly different best-fit parameters 
%%due to various systematic effects.
%%Moreover, none of these studies attempted an individual galaxy-by-galaxy comparison of their models 
%%with the previous literature. 
%%We have now made this comparison, identified the optimal decomposition, 
%%and obtained improved black hole mass scaling relations, 
%%which are detailed in this thesis. \\

Using \emph{Spitzer} observations at $3.6 ~\mu \rm m$, 
which is the best available wavelength band to trace the stellar mass, 
we performed state-of-the-art structural decompositions for $66$ galaxies 
with a direct measure of their black hole mass. 
Thanks to a meticulous inspection of each galaxy's substructure -- 
through photometric and isophotal analysis, unsharp masking, auxiliary information extracted from the literature, 
and, for the first time, kinematic maps -- 
we were able to identify \emph{a priori} the physical galaxy components. 
%Our multicomponent models account for spheroid, large-scale disc, 
%embedded or nuclear discs, spiral arms, bars, rings, halo, extended or unresolved nuclear source and partially depleted core. 
The combination of photometric and kinematic information was crucial 
to confirm the presence of rotationally supported components in most early-type (elliptical+lenticular) galaxies, 
and to identify their radial extent (nuclear, intermediate-scale or large-scale discs). 
Upon performing galaxy decompositions with both one-dimensional (1D) and two-dimensional (2D) parametric techniques, 
we observed no systematic differences between the results from 1D and 2D methods, 
but we found more advantages in the former. 
Our 66 galaxies constitute the largest sample to date 
for which these two conditions are satisfied: 
(\emph{i}) the spheroid structural parameters have been measured accurately and homogeneously;  
(\emph{ii}) and the black hole mass has been securely estimated with a direct method. 
%Unlike previous studies that mainly investigated massive early-type galaxies, 
%our sample contains a large number of spiral galaxies with low black hole masses ($<10^7 \rm~M_\odot$), 
%and allowed us to better investigate the poorly studied low-mass end of some black hole mass correlations. 
This thesis presents updates and modifications to several black hole mass scaling relations, 
and discusses important implications for galaxy evolution models. \\

mm + mn \\

ubermassive \\

>>>>

Supermassive black holes in local galaxies obey a surprisingly large number of scaling laws 
that involve the black hole mass and various properties of the host spheroid or galaxy. 
These ``black hole mass scaling relations'' reveal a strong symbiosis between galaxies and black holes, 
define important constraints about their co-evolution through the cosmic time, 
and set the boundary conditions (at $z=0$) for theoretical models and simulations of galaxy evolution.

Using \emph{Spitzer} observations at $3.6~\mu\rm m$, 
which is the best available wavelength band to trace the stellar mass, 
we performed state-of-the-art structural decompositions for $66$ galaxies 
with a direct measure of their black hole mass, $M_{\rm BH}$. 
Thanks to a meticulous inspection of each galaxy's substructure -- 
by means of photometric and isophotal analysis, unsharp masking, auxiliary information extracted from the literature, 
and, for the first time, kinematic maps -- 
we were able to identify \emph{a priori} the physical galaxy components. 
The combination of photometric and kinematic information was crucial 
to confirm the presence of rotationally supported components in most early-type (elliptical+lenticular) galaxies, 
and to identify their radial extent (nuclear, intermediate-scale or large-scale discs). 
Upon performing galaxy decompositions with both one-dimensional (1D) and two-dimensional (2D) parametric techniques, 
we observed no systematic differences between the results from 1D and 2D methods, 
but we found more advantages in the former. 
Our 66 galaxies constitute the largest sample to date 
for which the following two conditions are satisfied: 
(\emph{i}) the spheroid structural parameters have been measured accurately and homogeneously;  
(\emph{ii}) and the black hole mass has been securely estimated with a direct method. 
This thesis presents updates and modifications to several black hole mass scaling relations, 
and discusses important implications for galaxy evolution models.

For early-type (elliptical + lenticular) galaxies only, 
the black hole mass correlates equally well with galaxy luminosity, $L_{\rm gal}$, 
and spheroid luminosity, $L_{\rm sph}$, or spheroid stellar mass, $M_{\rm *,sph}$. 
However, when all galaxies (early- and late-type) are considered together, 
the $M_{\rm BH} - L_{\rm sph}$ relation has a lower level of intrinsic scatter than the $M_{\rm BH} - L_{\rm gal}$ relation. 
In the $M_{\rm BH} - M_{\rm *,sph}$ diagram, early-type galaxies define a tight linear correlation (a \emph{red sequence}), 
whereas late-type (spiral) galaxies follow a $2-3$ times steeper \emph{blue sequence}. 
The spheroid S\'ersic index, $n_{\rm sph}$, scales with $M_{\rm BH}$ 
in the same way for early- and late-type galaxies. 
Black holes that appear to be ``overmassive'' compared to expectations from the stellar velocity dispersion, $\sigma_*$, 
have been theoretically explained with their host galaxies having experienced more dry mergers than any other galaxy. 
In spite of that, we present empirical evidence supporting a scenario 
where the host galaxies of such overmassive black holes have undergone the lowest degree of dry merging. 
Finally, we debunk claims of four overmassive black holes in the $M_{\rm BH} - M_{\rm *,sph}$ diagram 
by demonstrating that the luminosity of their host spheroids had been considerably underestimated. 
We show that these four spheroids are unevolved relics of $z=2$ quiescent compact massive spheroids, 
and we confer about the significant consequences of this.

### 2016-03-22 thesis/conclusions.tex
corrections to chapters
thesis completed

\chapter{Final remarks and future perspectives}
\label{ch:concl}

@@

In Chapter \ref{ch:msigma}, we focused on the $M_{\rm BH} - \sigma$ diagram 
and in particular on the over-massive black holes that have been reported to exist 
at the high-mass end of the correlation. 
\citet{volontericiotti2013} explained the presence of the outlying black holes 
as a natural consequence of the fact that their host Central Cluster Galaxies (CCGs) have experienced more dry mergers 
than any other galaxy, 
since dry mergers are expected to increase the black hole mass 
and ``conserve'' the stellar velocity dispersion. 
We tested and disproved this scenario for 23 core-S\'ersic galaxies, 
using the ratio between black hole mass and stellar mass deficit as a proxy 
for the number of dry mergers experienced by a galaxy \citep{merritt2006}. 
This was confirmed by a second test using the central kinematic classification of 37 galaxies. 
Our analysis suggests that 
the merger history of the CCGs hosting over-massive black holes in the $M_{\rm BH} - \sigma$ diagram 
is not exceptional, 
therefore these galaxies should be considered as legitimate members of the $M_{\rm BH} - \sigma$ correlation. 
An alternative possibility for the CCGs is that their partially depleted cores 
have been replenished by stars, globular clusters and nuclear star clusters 
acquired from minor dry mergers, 
where the accreted satellite galaxies did not host a massive black hole. 
However, high-resolution cosmological simulations seem to favour an ``inside-out'' formation scenario 
for early-type galaxies, 
where minor mergers only contribute to the build-up of the external envelope of high-mass galaxies 
(e.g.~\citealt{wellons2016}). 
It should also be noted that evidence is accumulating against the ``naive'' assumption 
that low-mass galaxies do not host any central massive black hole 
(e.g.~\citealt{Baldassare2015,Graham2016LEDA}). \\

>>>>

\chapter{Final Remarks and Future Perspectives}
\label{ch:concl}

@@

In Chapter \ref{ch:msigma}, we focused on the $M_{\rm BH} - \sigma_*$ diagram 
and in particular on the over-massive black holes that have been reported to exist 
at the high-mass end of the correlation. 
\citet{volontericiotti2013} explained the presence of the outlying black holes 
as a natural consequence of the fact that their host Central Cluster Galaxies (CCGs) have experienced more dry mergers 
than any other galaxy, 
since dry mergers are expected to increase the black hole mass 
and ``conserve'' the stellar velocity dispersion. 
We tested and disproved this scenario for 23 core-S\'ersic galaxies, 
using the ratio between black hole mass and stellar mass deficit as a proxy 
for the number of dry mergers experienced by a galaxy \citep{merritt2006}. 
This was confirmed by a second test using the central kinematic classification of 37 galaxies. 
Our analysis suggests that 
the merger history of the CCGs hosting over-massive black holes in the $M_{\rm BH} - \sigma_*$ diagram 
is not exceptional, 
therefore these galaxies should be considered as legitimate members of the $M_{\rm BH} - \sigma_*$ correlation. 
An alternative possibility for the CCGs is that their partially depleted cores 
have been replenished by stars, globular clusters and nuclear star clusters 
acquired from minor dry mergers, 
where the accreted satellite galaxies did not host a massive black hole. 
However, high-resolution cosmological simulations seem to favour an ``inside-out'' formation scenario 
for early-type galaxies, 
where minor mergers only contribute to the build-up of the external envelope of high-mass galaxies 
(e.g.~\citealt{wellons2016}). 
It should also be noted that evidence is accumulating against the ``naive'' assumption 
that low-mass galaxies do not host any central massive black hole 
(e.g.~\citealt{Baldassare2015,Graham2016LEDA}). \\

### 2016-03-22 thesis/declaration.tex
acknowl.
thesis completed

I herewith declare that this thesis contains no material that has been accepted 
for the award to the candidate of any other degree or diploma. 
To the best of my knowledge, 
this thesis contains no material previously published or written by another author, 
except where due reference is made.
The work presented in this thesis has been carried out 
in the Centre for Astrophysics \& Supercomputing 
at Swinburne University of Technology between September 2012 and April 2016. 
The content of the chapters listed below has appeared in refereed journals.

@@

>>>>

I herewith declare that this thesis contains no material that has been accepted 
for the award to the candidate of any other degree or diploma. 
To the best of my knowledge, 
this thesis contains no material previously published or written by another author, 
except where due reference is made.
The work presented in this thesis has been carried out 
in the Centre for Astrophysics and Supercomputing 
at Swinburne University of Technology between September 2012 and April 2016. 
The content of the chapters listed below has appeared or will appear in refereed journals.

@@

\item Chapter \ref{ch:mn} has been accepted for publication in \emph{The Astrophysical Journal} as 
``Supermassive black holes and their host spheroids 
III. The $M_{\rm BH} - n_{\rm sph}$ correlation''
by Giulia A.~D.~Savorgnan.

### 2016-03-22 thesis/ellicular.tex
thesis completed

It has been claimed that a considerable number of high-redshift passive galaxies are significantly more compact 
than their local counterparts. 
\citet{daddi2005} reported on the observation of seven quiescent early-type galaxies 
at $z=1.4-2.5$ with stellar masses $\gtrsim 10^{11} \rm~M_\odot$ 
and effective radii significantly smaller than those of local counterparts. 
\citet{trujillo2006} analysed ten massive $\approx 5 \times 10^{11} \rm~M_\odot$ galaxies at $z=1.2-1.7$ 
and measured for them sizes a factor of 4 smaller than $z=0$ galaxies with similar stellar masses. 
From this, they concluded that the observed rapid evolution of the structural properties of massive quiescent galaxies 
over the last $\approx 10 \rm~Gyr$ 
cannot be reconciled with a monolithic formation scenario. 
\citet{kriek2008} and \citet{vandokkum2008} found that nearly half of $z \approx 2$ 
massive ($\approx 10^{11} \rm~M_\odot$) galaxies 
have old stellar populations, negligible star formation, and sizes a factor of 5 smaller than 
those of local descendants, 
and similar conclusions were reached by several other studies 
(e.g.\citealt{toft2007,trujillo2007,zirm2007,buitrago2008,damjanov2009}). 
While some studies have pointed out that ``progenitor bias''\footnote{``Progenitor bias'' refers to sample confusion 
of the local descendants of high-redshift massive quiescent galaxies. 
While some of the $z=0$ red-sequence galaxies might have had a passive evolution since $z \approx 2$, 
other $z=0$ red-sequence galaxies might descend from $z \approx 2$ star-forming galaxies. } 
(e.g.~\citealt{carollo2014}) might play an important role 
when tracking the evolution of massive quiescent galaxies, 
minor dry mergers have been commonly advocated to explain the size growth of these galaxies 
(e.g.~\citealt{hopkins2009,carrasco2010,cimatti2012,fan2013,de2014}).
However, not enough satellites have been found around massive galaxies 
to support the minor dry merger scenario 
(e.g.~\citealt{khochfarburkert2006,maller2006,hopkins2009,naab2009,mclure2013}). 
A fascinating solution to the problem was proposed by \citet{graham2013review}, 
who suggested that the high-redshift compact massive spheroids 
(also dubbed ``red nuggets'', \citealt{damjanov2009}) have evolved into the bulges 
of today's massive early-type disc galaxies (see also \citealt{dullograham2013cores} and \citealt{driver2013}). 
This view was confirmed by \citet{gds2015}, 
who used published, reliable bulge/disc decompositions 
to unveil a large number of previously unnoticed local stellar spheroidal systems 
with the same structural properties, old stellar populations, 
and similar number density as the high-redshift compact massive spheroids. 
\citet{gds2015} advocated the growth of two-dimensional stellar discs around the compact massive spheroids 
to explain the evolution of these objects. \\

>>>>

It has been claimed that a considerable number of high-redshift passive galaxies are significantly more compact 
than their local counterparts. 
\citet{daddi2005} reported on the observation of seven quiescent early-type galaxies 
at $z=1.4-2.5$ with stellar masses $\gtrsim 10^{11} \rm~M_\odot$ 
and effective radii significantly smaller than those of local counterparts. 
\citet{trujillo2006} analysed ten massive $\approx 5 \times 10^{11} \rm~M_\odot$ galaxies at $z=1.2-1.7$ 
and measured for them sizes a factor of 4 smaller than $z=0$ galaxies with similar stellar masses. 
From this, they concluded that the observed rapid evolution of the structural properties of massive quiescent galaxies 
over the last $\approx 10 \rm~Gyr$ 
cannot be reconciled with a monolithic formation scenario. 
\citet{kriek2008} and \citet{vandokkum2008} found that nearly half of $z \approx 2$ 
massive ($\approx 10^{11} \rm~M_\odot$) galaxies 
have old stellar populations, negligible star formation, and sizes a factor of 5 smaller than 
those of local descendants, 
and similar conclusions were reached by several other studies 
(e.g.\citealt{toft2007,trujillo2007,zirm2007,buitrago2008,damjanov2009}). 
While some studies have pointed out that ``progenitor bias''\footnote{``Progenitor bias'' refers to sample confusion 
of the local descendants of high-redshift massive quiescent galaxies. 
While some of the $z=0$ red-sequence galaxies might have had a passive evolution since $z \approx 2$, 
other $z=0$ red-sequence galaxies might descend from $z \approx 2$ star-forming galaxies. } 
(e.g.~\citealt{carollo2014}) might play an important role 
when tracking the evolution of massive quiescent galaxies, 
minor dry mergers have been commonly advocated to explain the size growth of these galaxies 
(e.g.~\citealt{hopkins2009,carrasco2010,cimatti2012,fan2013,de2014}).
However, not enough satellites have been found around massive galaxies 
to support the minor dry merger scenario 
(e.g.~\citealt{khochfarburkert2006,maller2006,hopkins2009,naab2009,mclure2013}). 
A fascinating solution to the problem was proposed by \citet{graham2013review}, 
who suggested that the high-redshift compact massive spheroids 
(also dubbed ``red nuggets'', \citealt{damjanov2009}) have evolved into the bulges 
of today's massive early-type disc galaxies (see also \citealt{dullograham2014cores} and \citealt{driver2013}). 
This view was confirmed by \citet{gds2015}, 
who used published, reliable bulge/disc decompositions 
to unveil a large number of previously unnoticed local stellar spheroidal systems 
with the same structural properties, old stellar populations, 
and similar number density as the high-redshift compact massive spheroids. 
\citet{gds2015} advocated the growth of two-dimensional stellar discs around the compact massive spheroids 
to explain the evolution of these objects. \\

### 2016-03-22 thesis/introduction.tex
thesis completed

The discovery of ultraluminous quasars at $z>6$ has accentuated the urge to create very massive black hole seeds in a relatively short time 
(e.g.~\citealt{alexandernatarajan2014,madau2014,lupi2015}). 
To date, there have been nearly 50 claims of $z>6$ quasars hosting $\gtrsim 10^9~\rm M_\odot$ black holes 
(e.g.~\citealt{fan2003,jiang2007,mortlock2011,banados2014,trakhtenbrot2015,wu2015}). 
Within a $\Lambda$CDM cosmology, such early giant monsters could not have formed so quickly 
without the creation of anomalously massive seeds 
or incredibly high accretion rates, i.e.~exceeding the Eddington limit 
(but see \citealt{meliamcclintock2015} for an alternative explanation). 
However, it is worth noting that these black hole masses are calculated with the reverberation mapping method, 
assuming that $M_{\rm BH}$ is directly proportional to the virial factor $f$, 
which is calibrated on the normalisation of the $z=0$ observed $M_{\rm BH} - \sigma_*$ correlation. 
The existence of (unknown) selection biases in the local sample of directly measured black hole masses 
would imply a systematic overestimation of the virial factor and, consequently, 
of the black hole masses estimated with the reverberation mapping technique. \\

@@

\cite{vandenbosch2012} combined archival Hubble Space Telescope (\emph{HST}) high-resolution imaging 
and long-slit spectroscopy (obtained with the Marcario Low Resolution Spectrograph on the Hobby-Eberly Telescope, Texas) 
of the nuclear region of the galaxy NGC 1277, 
and directly measured the central black hole mass by fitting self-consistent Schwarzschild models 
to the integrated stellar kinematics. 
They found a twenty billion solar masses black hole, 
which they estimated to weight 59\% of the host spheroid's mass (or 14\% of the galaxy's total stellar mass). 
According to their measurement, 
the black hole of NGC 1277 was an order of magnitude more massive than what expected from the host spheroid's stellar velocity dispersion 
and two orders of magnitude more massive than what expected from the host spheroid's stellar mass.
\cite{emsellem2013} re-analysed \citeauthor{vandenbosch2012}'s data 
and showed that their observations were consistent with a black hole mass up to an order of magnitude smaller 
than the extraordinary value previously reported. 
However, \cite{yildirim2015} and \cite{scharwachter2015} confirmed \citeauthor{vandenbosch2012}'s measurement. \\

>>>>

The discovery of ultraluminous quasars at $z>6$ has accentuated the urge to create very massive black hole seeds in a relatively short time 
(e.g.~\citealt{alexandernatarajan2014,madau2014,lupi2016}). 
To date, there have been nearly 50 claims of $z>6$ quasars hosting $\gtrsim 10^9~\rm M_\odot$ black holes 
(e.g.~\citealt{fan2003,jiang2007,mortlock2011,banados2014,trakhtenbrot2015,wu2015}). 
Within a $\Lambda$CDM cosmology, such early giant monsters could not have formed so quickly 
without the creation of anomalously massive seeds 
or incredibly high accretion rates, i.e.~exceeding the Eddington limit 
(but see \citealt{meliamcclintock2015} for an alternative explanation). 
However, it is worth noting that these black hole masses are calculated with the reverberation mapping method, 
assuming that $M_{\rm BH}$ is directly proportional to the virial factor $f$, 
which is calibrated on the normalisation of the $z=0$ observed $M_{\rm BH} - \sigma_*$ correlation. 
The existence of (unknown) selection biases in the local sample of directly measured black hole masses 
would imply a systematic overestimation of the virial factor and, consequently, 
of the black hole masses estimated with the reverberation mapping technique. \\

@@

\cite{vandenbosch2012} combined archival Hubble Space Telescope (\emph{HST}) high-resolution imaging 
and long-slit spectroscopy (obtained with the Marcario Low Resolution Spectrograph on the Hobby-Eberly Telescope, Texas) 
of the nuclear region of the galaxy NGC 1277, 
and directly measured the central black hole mass by fitting self-consistent Schwarzschild models 
to the integrated stellar kinematics. 
They found a twenty billion solar masses black hole, 
which they estimated to weight 59\% of the host spheroid's mass (or 14\% of the galaxy's total stellar mass). 
According to their measurement, 
the black hole of NGC 1277 was an order of magnitude more massive than what expected from the host spheroid's stellar velocity dispersion 
and two orders of magnitude more massive than what expected from the host spheroid's stellar mass.
\cite{emsellem2013} re-analysed \citeauthor{vandenbosch2012}'s data 
and showed that their observations were consistent with a black hole mass up to an order of magnitude smaller 
than the extraordinary value previously reported. 
However, \cite{yildirim2015} and \cite{scharwachter2016} confirmed \citeauthor{vandenbosch2012}'s measurement. \\
